---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# tidylog

The goal of tidylog is to provide feedback about basic dplyr operations. It provides simple wrapper functions for the most common functions, such as `filter`, `mutate`, `select`, and `group_by`.

## Example

Load `tidylog` after `dplyr`:

```{r message=FALSE}
library("dplyr")
library("tidylog")
```

Tidylog will give you feedback, for instance when filtering a data frame:

```{r}
filtered <- filter(mtcars, cyl == 4)
```

This can be especially helpful in longer pipes:

```{r}
summary <- mtcars %>%
    select(mpg, cyl, hp) %>%
    filter(mpg > 15) %>%
    mutate(mpg_round = round(mpg)) %>%
    mutate(mpg_round = ifelse(mpg_round > 30, NA, mpg_round)) %>%
    filter(!is.na(mpg_round)) %>%
    group_by(cyl, mpg_round) %>%
    tally() %>%
    filter(n >= 1)
```
Here, it might have been accidental that the last `filter` command had no effect.

## Installation

``` r
devtools::install_github("elbersb/tidylog")
```
